---
title: "GAMM Networks"
author: "Nick Gauthier"
date: "February 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
GAMMs are a flexible form of regression model well-matched to the complexities of the archaeological record, including non-normal distributions in the form of counts or proportions, non-independent observations with correlated errors, and non-linear functional relationships. Using two case studies -- an ethnographic marriage network and an archaeological assemblage similarity network -- I illustrate how this approach can lead to unbiased parameter estimates and more robust comparisons of competing hypotheses. I conclude by outlining how future empirical efforts can help to refine our thinking about past network dynamics and reveal cross-cultural regularities of human social interaction in the present day.

# Case Studies

Distance is a fundamental constraint on human social interaction. This basic principle motivates the use of spatial interaction models for estimating flows of people, information, and resources on spatial and social networks. These models have both valid dynamical and statistical interpretations, a key strength well supported by theory and data from geography, economics, ecology, and genetics. To date, archaeologists have primarily relied on the dynamical approach because the idiosyncrasies of archaeological data make the wholesale adoption of statistical approaches from other fields impractical.
# Setup

```{r }
library(raster)
library(archdata)
library(maps)
library(tidyverse)
library(readxl)
library(mgcv)
library(ggmap)
library(sf)
# devtools::install_github('nspope/corMLPE')
library(corMLPE)
library(tidygraph)
library(ggraph)
```

## Oxford Pots
```{r}
data("OxfordPots")

pots <- OxfordPots %>%
  rename(to = Place) %>%
  gather(key, value, OxfordPct:NewForestDst) %>%
  separate(key, c('from', 'measure'), sep = -3) %>%
  spread(measure, value) %>%
  rename(percent = Pct, distance = Dst) %>%
  mutate(from = str_replace(from, 'NewForest', 'New Forest'))

pots
```

```{r}
ggplot(pots, aes(distance, percent, color = from)) +
  geom_point(aes(shape = as.factor(WaterTrans)))
```

```{r, eval = FALSE}
ggmap::register_google(key = "<your-api-key-here>")

pots_sites <- pots %>%
  select(from, to) %>% 
  gather %>%
  pull(value) %>%
  unique %>% 
  tibble(site = .) %>%
  mutate(search_term = str_replace_all(site, c('Clausentum' = 'Bitterne', 'Mildenhall' = 'Mildenhall, Wiltshire')),
         search_term = paste0(search_term, ', UK')) %>%
  mutate_geocode(search_term) %>%
  select(-search_term) %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>%
  bind_cols(., as_tibble(st_coordinates(.))) %>%
  rename(x = X, y = Y)

saveRDS(pots_sites, 'oxfordpots_locations.RDS')
```

```{r}
pots_sites <- readRDS('oxfordpots_locations.RDS') %>%
  left_join(select(pots, to, WalledArea) %>% group_by(to) %>% 
              summarise(area = mean(WalledArea)), by = c('site' = 'to'))
```

```{r, eval = FALSE}
st_distance(pots_sites) %>%
  `colnames<-`(pots_sites$site) %>%
  as_tibble %>%
  mutate(to = pots_sites$site) %>%
  gather(from, distance, `New Forest`:Wroxeter) %>%
  filter(to > from) %>%
  right_join(dat, by = c('from' = 'origin', 'to' = 'destination')) %>%
  mutate(distance.x = as.numeric(distance.x) / 1000) %>%
  ggplot(aes(distance.y, distance.x, color = from)) +
  geom_point() +
  geom_label(aes(label = to), alpha = .4)
```

```{r}
pots_net <- tbl_graph(nodes = pots_sites, edges = pots) %E>%
  filter(!is.na(percent)) %>%
  mutate(similarity = percent / 100)
```

```{r}
uk_boundary <- maps::map('world', region = 'UK',
                    fill = TRUE, plot = FALSE) %>%
  st_as_sf
```

```{r}
pots_net %E>%
  arrange(percent) %>%
ggraph() +
  geom_sf(data = uk_boundary, fill = NA) +
  geom_edge_link(aes(colour = percent), edge_width = 2, lineend = 'round') +
  geom_sf(data = filter(pots_sites, site %in% c('Oxford', 'New Forest')), size = 4) +
  scale_edge_color_distiller(palette = 'YlGnBu', guide = 'legend', name = 'Percentage of pottery \nfrom origin site') +
  labs(title = 'Ceramic distribution in Roman Britain', subtitle = 'Percentages of late Romano-British pottery produced in Oxford and New Forest') +
  coord_sf(datum = NA, xlim = c(-6, 2), ylim = c(50, 53.7)) +
  theme_void()
```

```{r}
pots_dat <- pots_net %E>%
  mutate(water = as.factor(WaterTrans == 1),
         x = .N()$x[to],
         y = .N()$y[to]) %>% 
  as_tibble %>%
  mutate(from = as.factor(from),
         to = as.factor(to))
```

Example of model selection.
```{r}
pots_models <- c("s(distance) + from",
            "s(distance) + water + from",
            "s(distance, by = water) + water + from") %>%
  paste('similarity ~', .) %>%
  tibble(formula_chr = .) %>%
  mutate(formula = map(formula_chr, as.formula), 
         model = map(formula, ~gam(.x, method = 'ML', family = betar(), data = pots_dat)),
         aic = map_dbl(model, AIC)) %>%
  arrange(aic)

pots_models
```

Refit the best performing model with REML and investigate.
```{r}
m1 <- gam(similarity ~ s(distance, by = water) + water + from,
          method = 'REML',
          select = TRUE,
          family = betar(),
          data = pots_dat)

summary(m1)
plot(m1, trans = plogis)
```
```{r echo = FALSE}
tmp1 <- expand.grid(distance = 0:140, water = c(TRUE, FALSE), from = 2)
predict(m1, tmp1, type = 'response', se.fit = TRUE) %>%
  bind_cols(tmp1) %>%
  mutate(fit = fit * 100,
         lower = fit - 2 * se.fit * 100,
         upper = fit + 2 * se.fit * 100) %>%
  ggplot(aes(distance, fit, group = water)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = water), alpha = .2) +
  geom_line(aes(y = fit, color = water), size = 1.2) +
  theme_minimal() +
  labs(title = 'Estimated distance decay functions', subtitle = 'With and without water transport',
       x = 'Distance (km)', y = 'Percentage of pottery from origin site')
```

```{r}
pots_net %E>%
  mutate(res = residuals(m1, type = 'response') * 100) %>%
  arrange(abs(res)) %>%
ggraph() +
  geom_sf(data = uk_boundary, fill = NA) +
  geom_edge_link(aes(colour = res), edge_width = 2, lineend = 'round') +
  geom_sf(data = filter(pots_sites, site %in% c('Oxford', 'New Forest')), size = 4) +
  scale_edge_color_distiller(palette = 'RdBu', limits = c(-8, 8), labels = paste0(seq(-8, 8, by = 4), '%'), guide = 'legend', name = '') +
  labs(title = 'Residuals from a distance-only interaction model', subtitle = 'Red ties are stronger then expected by distance, blue ties weaker') +
  coord_sf(datum = NA, xlim = c(-6, 2), ylim = c(50, 53.8)) +
  theme_void()
```

## Data Import and Preprocessing

```{r}
chumash <- read.csv('chumash_marriage.csv', row.names = 1) %>%
  as.matrix %>% 
  replace(. == 0, 999) %>% # replace 0 values with 999 temporarily
  as_tbl_graph(directed = FALSE) %E>%
  filter(!edge_is_loop()) %>%
  rename(marriages = weight) %>%
  mutate(marriages = if_else(marriages == 999, 0, marriages))%N>%
  mutate(centrality = centrality_degree(weights = marriages))
```

```{r}
chumash_locs <- read_csv('chumash_villages_locations.csv') %>%
  st_as_sf(coords = c('X', 'Y'), crs = 4326) %>%
  bind_cols(., as_tibble(st_coordinates(.))) %>%
  rename(x = X, y = Y) %>%
  mutate(size = ordered(size, levels = c('hamlet', 'small', 'medium', 'large', 'very large', 'largest')),
         population = case_when(size == 'hamlet' ~ 15,
          size == 'small' ~ 30,
          size == 'medium' ~ 60,
          size == 'large' ~ 120,
          size == 'very large' ~ 200,
          size == 'largest' ~ 300
          )) 

chumash_net <- left_join(chumash, chumash_locs, by = c('name' = 'village')) %E>%
  mutate(distance = st_distance(.N()$geometry[from], .N()$geometry[to], by_element = TRUE),
         distance = as.numeric(distance) / 1000) %>%
  mutate(eco= case_when(.N()$coastal[from] == .N()$coastal[to] & .N()$coastal[from] == TRUE ~ 'Coastal',
           .N()$coastal[from] == .N()$coastal[to] & .N()$coastal[from] == FALSE ~ 'Inland',
           .N()$coastal[from] != .N()$coastal[to] ~ 'Cross Boundary'),
         eco = as.factor(eco),
         eco_same = as.factor(.N()$coastal[from] == .N()$coastal[to]))

ca_boundary <- maps::map('county', region = 'California',
                    fill = TRUE, plot = FALSE) %>%
  st_as_sf

```

```{r}
chumash_net %E>%
  arrange(marriages) %>%
  filter(marriages > 0) %>%
ggraph() +
  geom_edge_link(aes(colour = marriages, alpha= marriages, width = marriages), lineend = 'round') +
  geom_node_point(aes(size = population, color = coastal)) +
  scale_edge_alpha(range = c(.5, 1)) +
  scale_edge_color_distiller(palette = 'YlGnBu') + 
  theme_void() +
  scale_size_area() +
  coord_quickmap() +
  geom_sf(data = ca_boundary, fill = NA) +
  coord_sf(datum = NA, xlim = c(-120.60246, -119.48277), ylim = c(34.38628, 34.74438 ))

```
468 marriages
```{r}

test <- seq(1.10, 1.30, .01) %>%
  map(~gamm(marriages ~ s(distance, by = eco) + eco,
      method = 'REML',
      select = TRUE,
      family =  Tweedie(p = .x, link = 'log'), #quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges'))))

test2 <- seq(1.10, 1.30, .01) %>%
  map(~gamm(marriages ~ s(distance, by = eco_same) + eco_same,
      method = 'REML',
      select = TRUE,
      family =  Tweedie(p = .x, link = 'log'), #quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges'))))
  
map(test, ~(.$lme) %>% AIC)
map(test2, ~(.$lme) %>% AIC) %>% unlist %>% plot

m1 <- gamm(marriages ~ s(distance, by = eco) + eco,
      method = 'REML',
      select = TRUE,
      family =  Tweedie(p = 1.2, link = 'log'), #quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges')))
m2 <- gamm(marriages ~ s(distance, by = eco) + eco,
      method = 'REML',
      select = TRUE,
      family =  quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges')))
AIC(m2)

plot(m1$gam, trans = exp)
gratia::draw(m1$gam)
summary(m1$gam)
summary(m1$lme)
gam.check(m1$gam)

chumash_net %E>%
  as_tibble %>%
  ggplot(aes(distance, marriages)) +
  geom_point() +
  geom_smooth()
```

```{r echo = FALSE}
tmp1 <- expand.grid(distance = 0:60, eco = c('Coastal', 'Cross Boundary', 'Inland')) #= c('TRUE', 'FALSE'))
predict(m1$gam, tmp1, type = 'response', se.fit = TRUE) %>%
  bind_cols(tmp1) %>%
  mutate(fit = fit,
         lower = fit - 2 * se.fit,
         upper = fit + 2 * se.fit) %>%
  ggplot(aes(distance, fit, group = eco)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = eco), alpha = .2) +
  geom_line(aes(y = fit, color = eco), size = 1.2) +
  theme_classic() +
  labs(title = 'Estimated distance decay functions', subtitle = 'Within and across ecological boundaries',
       x = 'Distance (km)', y = 'Number of marriages')
```

```{r}
tmp2 <- match(predict(m1$gam), predict(m1$lme))

predict(m1$gam) == predict(m1$lme)

chumash_net %E>%
  mutate(pred = predict(m1$lme),
         pred2 = predict(m1$gam, type = 'response'),
         res = marriages - exp(pred)) %>%
  #mutate(res = residuals(m1$gam, type = 'response')) %>%
  arrange(abs(res)) %>%
  filter(marriages != 0) %>%
ggraph() +
  geom_edge_link(aes(colour = res, alpha= res, width = res), lineend = 'round') +
  geom_node_point(aes(size = population, color = coastal)) +
  scale_edge_alpha(range = c(.5, 1)) +
  scale_edge_color_distiller(palette = 'RdBu', limits = c(-10, 10)) + 
  theme_void() +
  scale_size_area() +
  coord_quickmap() +
  geom_sf(data = ca_boundary, fill = NA) +
  coord_sf(datum = NA, xlim = c(-120.60246, -119.48277), ylim = c(34.38628, 34.74438 ))
```



```{r}
m1 <- gamm(qlogis(similarity) ~ s(distance, bs = 'cr') + as.factor(time), 
           data = dat, 
           correlation = corMLPE(form = ~from + to|as.factor(time)), 
           method = 'ML')
m2 <- gamm(qlogis(similarity) ~ s(distance, bs = 'cr', by = as.factor(time)) + as.factor(time), 
           data = dat, 
           correlation = corMLPE(form = ~from + to|as.factor(time)), 
           method = 'ML')
m3 <- gamm(qlogis(similarity) ~ s(distance, bs = 'cr', m = 2) + s(distance, bs = 'cs', by = as.factor(time), m = 1) +
             as.factor(time), 
           data = dat, 
           correlation = corMLPE(form = ~from + to|as.factor(time)), 
           method = 'ML')
AIC(m1$lme, m2$lme, m3$lme)

plot(m1$gam, trans = plogis)
plot(m2$gam, trans = plogis)
plot(m3$gam, trans = plogis)


AIC(m3$lme, m4$lme)
# moderate support for regional effect
summary(m1$gam)
summary(m2$gam)
summary(m3$lme)


ggplot(dat, aes(distance)) + geom_histogram()
ggplot(dat, aes((similarity))) + geom_histogram()
ggplot(dat, aes(qlogis(similarity), fill = time)) + geom_histogram()
sum(min(dat$similarity))

gam.check(m3$gam)
```

```{r}
mod_time <- dat %>%
  group_by(time) %>%
  nest %>%
  mutate(model = map(data, ~gamm(qlogis(similarity) ~ s(distance, bs = 'cr'), 
                                 data = ., correlation = corMLPE(form = ~from + to), method = 'ML')),
         r2 = map_dbl(model, ~summary(.$gam)$r.sq))
mod_time

mod_time$model %>%
  map(~.$gam) %>%
  walk(plot, trans = plogis)
```


