---
#---POSTER SIZE & DEFAULT FONT---#
poster_height: "36in" # height in inches of poster
poster_width: "48in" # width in inches of poster
font_family: "palatino" # choose from typical latex fonts (example: "palatino")
font_size: "30pt" #please see github.com/brentthorne/posterdown for compatible options.

#---TITLE BOX OPTIONS---#
#ESSENTIALS
title: '\fontfamily{phv}\selectfont Generalized additive mixed models for spatial networks'
author: "Nicolas Gauthier"
affiliation: "School of Human Evolution and Social Change, Arizona State University"
#STYLE & FORMATTING
titlebox_bgcol: "433E85"  #Colour of the Title Box background
titlebox_bordercol: "433E85" #Colour of the title Box border.
titlebox_shape: "all"
titlebox_borderwidth: "1cm"
title_textcol: "FFFFFF" #colour of title text
author_textcol: "ffffff" # Colour of author text
affiliation_textcol: "ffffff" # Colour of affiliation text
title_textsize: "Huge"         # Poster title fontsize
author_textsize: "Large"       # Author list font size
affiliation_textsize: "large"  # Affiliation font size
#ADDING LOGOS
logoleft_name: 'figures/shesc'
logoleft_width: '5in'
logoleft_xshift: '1in'
logoleft_yshift: '1in'
logoright_name: 'figures/quantarch3'
logoright_width: '5in'
logoright_xshift: '-1in'
logoright_yshift: '1in'

#---POSTER BODY OPTIONS---#
body_bgcol: "FFFFFF" #colour of the poster main background
body_textsize: "normalsize"    # Size of the main poster body text
body_textcol: "594F4F"#000000" # Colour of main text in the body of poster
column_numbers: 3 # Number of columns that the poster has
column_margins: "0.5in" # Margin spacing for columns
columnline_col: "008080" #colour 
columnline_width: "0pt" #width of line between each column
#SECTION TITLE STYLING
sectitle_textcol: "ffffff" # Colour of the poster section titles
sectitle_bgcol: "433E85" # Colour of the section title box
sectitle_bordercol: "433E85" # Colour of the border around the section title box.
sectitle_borderwidth: "2mm" # Thicknes of the section title box border
sectitle_boxshape: "uphill" # Changes the shape of the section title box.

#---BIBLIOGRAPHY OPTIONS---#
bibliography: references # name of the .bib file used for referencing
bibliography_spacing: 0.8 # sets the multiplier for line spacing of bibliography spacing (between 0 and 1)
bibliography_textsize: "small"  # size of the bibliography text size (handy for one too many references!)

#---OTHER---#
cite_col: "CC0000" #colour of ciation elements
url_col: "008080" #colour of url links
link_col: "008080" #colour of other links within the poster
footnote_textcol: "000000" # Colour of footnote text if used
output: posterdown::posterdown_latex
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis', out.width = "80%", warning = FALSE, fig.align = 'center', message = FALSE, dev='cairo_pdf', cache = TRUE)
```

```{r}
#devtools::install_github('brentthorne/posterdown')
#library(posterdown)
library(raster)
library(ggspatial)
library(maps)
library(tidyverse)
library(mgcv)
library(sf)
#devtools::install_github('nspope/corMLPE')
library(corMLPE)
library(tidygraph)

# need dev version of ggraph 
# devtools::install_github('thomasp85/ggraph')
library(ggraph)
library(archdata)

library(patchwork)
```
# What are GAMMs?

Generalized additive models (GAMs) and their variants Generalized additive *mixed* models (GAMMs) are a flexible family of regression models that capture the complexity of the archaeological record.

\vspace{1cm}

**Let's break it down:**

* *Generalized* -- Model non-normal distributions such as counts or proportions
* *Additive* -- Combine linear and non-linear relationships
* *Mixed* -- Control for network, spatial, and temporal autocorrelation

# What can I do with them?

In archaeology, they're particuarly useful for fitting so-called ``maximum-entropy'' spatial interaction models. These models estimate the *flow* of goods, information, or people between spatially-structured populations as a function of the origin site, destination site, and the space between them:

\vspace{1cm}

\begin{center}
  \includegraphics[width=.4\columnwidth]{figures/network_diagram.png}
\end{center}
$$flow = f\left(origin\right) \times f\left(destination\right) \times f\left(distance\right)$$

The difficult part comes when we have to define the $f\left(\right)$s. Economic geographers often use a generalized *linear* model (GLM), which requires them to definte $f\left(\right)$ ahead of time. In a GAM, the $f\left(\right)$s are estimated directly from the data using splines.

\vspace{1cm}

# How do they work?

\vspace{1cm}

\begin{center}
\includegraphics[width=.3\columnwidth]{figures/spline.png}
\end{center}

\vspace{1cm}

Real-world splines are flexible strips of metal or wood used to draw curves. Mathematical splines are complex curves made of many smaller, simpler curves. *Penalized* regression splines can estimate $f\left(\right)$ from the data, limiting overfitting by penalizing the "wiggliness" of the function.



\vspace{1cm}

# Case Studies

```{r, include = FALSE}
data("OxfordPots")
pots <- OxfordPots %>%
  rename(to = Place) %>%
  gather(key, value, OxfordPct:NewForestDst) %>%
  separate(key, c('from', 'measure'), sep = -3) %>%
  spread(measure, value) %>%
  rename(percent = Pct, distance = Dst) %>%
  mutate(from = str_replace(from, 'NewForest', 'New Forest'))

pots_sites <- readRDS('data/oxfordpots_locations.RDS') %>%
  left_join(dplyr::select(pots, to, WalledArea) %>% group_by(to) %>% 
              summarise(area = mean(WalledArea)), by = c('site' = 'to'))

pots_net <- tbl_graph(nodes = pots_sites, edges = pots, directed = TRUE) %E>%
  filter(!is.na(percent)) %>%
  mutate(similarity = percent / 100)

uk_boundary <- maps::map('world', region = 'UK',
                    fill = TRUE, plot = FALSE) %>%
  st_as_sf

pots_dat <- pots_net %E>%
  mutate(water = as.factor(WaterTrans == 1),
         x = .N()$x[to],
         y = .N()$y[to]) %>% 
  as_tibble %>%
  mutate(from = as.factor(from),
         to = as.factor(to))
```

```{r, out.width ="100%"}
pots_net %E>%
  arrange(percent) %N>%
ggraph(., layout = 'manual', x = as_tibble(.)$x, y = as_tibble(.)$y) +
  geom_sf(data = uk_boundary, fill = NA) +
  geom_edge_link(aes(colour = percent), edge_width = 2, lineend = 'round') +
  geom_sf(data = filter(pots_sites, site %in% c('Oxford', 'New Forest')), size = 4) +
  scale_edge_color_viridis(guide = 'legend', name = 'Pottery \nfrom origin site',
                           labels = paste0(seq(5, 20, 5), '%'),
                           breaks = seq(5, 20, 5)) +
  annotation_scale(location = 'br', style = 'ticks') +
  coord_sf(datum = NA, xlim = c(-6, 2), ylim = c(50, 53.8)) +
  ggtitle('Ceramic distribution in Roman Britain', 
          'Percent late Romano-British pottery produced in Oxford and New Forest \n \n \n') +
  theme_void()
```
The Oxford Pots dataset can be found in the R package `archdata`. \emph{\small{David L. Carlson and Georg Roth (2018). archdata: Example Datasets from Archaeological Research. R package version 1.2.}}

\vspace{4cm}


```{r}
chumash_villages <- read_csv('data/chumash_villages.csv') %>%
  st_as_sf(coords = c('x', 'y'), crs = 4326) %>%
  bind_cols(., as_tibble(st_coordinates(.))) %>%
  rename(x = X, y = Y) %>%
  mutate(size = ordered(size, levels = c('hamlet', 'small', 'medium', 'large', 'very large', 'largest'))) 

chumash_net <- read.csv('data/chumash_marriages.csv', row.names = 1) %>%
  as.matrix %>% 
  replace(. == 0, 999) %>% # replace 0 values with 999 temporarily
  as_tbl_graph(directed = FALSE) %E>%
  filter(!edge_is_loop()) %>%
  rename(marriages = weight) %>%
  mutate(marriages = if_else(marriages == 999, 0, marriages))%N>%
  mutate(centrality = centrality_degree(weights = marriages)) %>%
  left_join(chumash_villages, by = c('name' = 'village')) %E>%
  mutate(distance = st_distance(.N()$geometry[from], .N()$geometry[to], by_element = TRUE),
         distance = as.numeric(distance) / 1000,
         baptisms = .N()$baptisms[from] * .N()$baptisms[to],
         eco_same = as.factor(.N()$coastal[from] == .N()$coastal[to]))

chumash_dat <-  chumash_net %E>% as_tibble

elev <- raster('data/santa_barbara_dem.tif') %>% 
  as.data.frame(xy = TRUE, na.rm = TRUE)

chumash_paths <- read_sf('data/chumash_paths.shp')
```

```{r, out.width ="100%"}
chumash_paths %>%  
  arrange(marriages) %>%
  ggplot +
  geom_raster(data = elev, aes(x, y, fill = santa_barbara_dem), show.legend = FALSE) +
  geom_sf(aes(color = marriages), size = 2) +
  geom_sf(data = chumash_villages, aes(size = baptisms), show.legend = 'point') +
  scale_size_area(name = 'Baptisms') +
  scale_color_viridis_c(name = '    Marriages\n ') +
  annotation_scale(style = 'ticks', pad_x = unit(4.35, 'cm'), pad_y = unit(.5, 'cm')) +
  scale_fill_viridis_c(option = 'D', alpha = .6) +
  theme_void() +
  coord_sf(datum = NA) +
  ggtitle('Coastal Chumash marriages in southern California', 
          'Marriages in mission registers, population estimated from baptisms') +
  theme(legend.position = 'bottom')
```
Chumash marriage data can be found in John Johnson's PhD thesis. \emph{\small{John R. Johnson, “Chumash social organization: An ethnohistorical perspective”. PhD thesis. UC Santa Barbara, 1988.}}

# Results
## Late Romano-British Pottery
```{r}
m1 <- gam(similarity ~ s(distance, by = water) + water + from,
          family = betar(), # beta distribution between 0 and 1
          method = 'REML', # algorithm to estimate the GAM
          select = TRUE, # allow variables to be selected out
          data = pots_dat)
```

```{r, out.wdith = "55%"}
tmp1 <- expand.grid(distance = 0:140, water = c(TRUE, FALSE), from = 2)
predict(m1, tmp1, type = 'response', se.fit = TRUE) %>%
  bind_cols(tmp1) %>%
  mutate(fit = fit * 100,
         lower = fit - 2 * se.fit * 100,
         upper = fit + 2 * se.fit * 100) %>%
  ggplot(aes(distance, fit, group = water)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = water), alpha = .2) +
  geom_line(aes(y = fit, color = water), size = 1.2) +
  scale_color_manual(values = c('#22A884', '#414487'), name = 'Transport', labels = c('Land', 'Water')) +
  scale_fill_manual(values = c('#22A884', '#414487'), name = 'Transport', labels = c('Land', 'Water')) +
  theme_classic() +
  labs(x = 'Distance (km)', y = 'Pottery from origin site (%)',
       title = 'Distance decay functions on land and water') 

# pots_net %E>%
#   mutate(res = residuals(m1, type = 'response') * 100) %>%
#   arrange(abs(res)) %N>%
# ggraph(., layout = 'manual', x = as_tibble(.)$x, y = as_tibble(.)$y) +
#   geom_sf(data = uk_boundary, fill = NA) +
#   geom_edge_link(aes(colour = res), edge_width = 2, lineend = 'round') +
#   geom_sf(data = filter(pots_sites, site %in% c('Oxford', 'New Forest')), size = 4) +
#   scale_edge_color_distiller(palette = 'RdBu', limits = c(-8, 8), labels = paste0(seq(-8, 8, by = 4), '%'), guide = 'legend', name = 'Residuals') +
#   coord_sf(datum = NA, xlim = c(-6, 2), ylim = c(50, 53.8)) + 
#   labs(caption = 'Red ties stronger then expected, blue ties weaker') +
#   theme_void() 
```

GAMs learn functions directly from data. Here we find that the amount of pottery moved over land decays logarithmically, but pottery moved by water does not.

## Mission-period Chumash

```{r, include=FALSE}
m2 <- gamm(marriages ~ te(distance, baptisms) + eco_same,
           family =  quasipoisson(),
           correlation = corMLPE(form = ~from + to),
           method = 'REML',
           select = TRUE,
           data = chumash_dat)
```

```{r}
p1 <- expand.grid(baptisms = seq(0, 300)^2,
            distance = seq(0, 30, 0.1),
            eco_same = FALSE) %>%
  mutate(marriages = predict(m2$gam, ., type = 'response')) %>%
  ggplot(aes(distance, sqrt(baptisms), fill = marriages)) +
  geom_raster() +
  scale_fill_viridis_c(guide = 'legend', breaks = c(0,2,4,6,8,10,12), name = 'Marriages') +
  labs(title = 'Tensor spline of population and distance', x = 'Distance (km)', y = 'Baptisms') +
  theme_classic() +
  coord_fixed(ratio = .1)

ggsave('figures/tensor_2d.pdf', plot = p1)
```

\begin{center}
\includegraphics[width=.49\columnwidth]{figures/tensor_2d.pdf}
\includegraphics[width=.39\columnwidth]{figures/tensor_3d.pdf}
\end{center}

Here we can use a ``tensor'' spline to model the nonlinear influence of distance and population size on Chumash marriage patterns.

# What's next?

Efficient estimation of GAMs and GAMMs is available in `R` with the package `mgcv`, and `brms` provides a Bayesian implementation. Scan this QR code or go to https://github.com/nick-gauthier/gam-networks for the code and data for this poster along with more detailed worked examples.  

\begin{center}
\includegraphics[width=3.3in]{figures/qr}
\end{center}
