---
title: "Generalized additive mixed models for spatial networks"
author: "Nick Gauthier"
output: 
  html_document: 
    highlight: zenburn
    keep_md: yes
    theme: cosmo
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
GAMMs are a flexible form of regression model well-matched to the complexities of the archaeological record, including non-normal distributions in the form of counts or proportions, non-independent observations with correlated errors, and non-linear functional relationships. Using two case studies -- an ethnographic marriage network and an archaeological assemblage similarity network -- I illustrate how this approach can lead to unbiased parameter estimates and more robust comparisons of competing hypotheses.

Distance is a fundamental constraint on human social interaction. This basic principle motivates the use of spatial interaction models for estimating flows of people, information, and resources on spatial and social networks. These models have both valid dynamical and statistical interpretations, a key strength well supported by theory and data from geography, economics, ecology, and genetics. To date, archaeologists have primarily relied on the dynamical approach because the idiosyncrasies of archaeological data make the wholesale adoption of statistical approaches from other fields impractical.

# Setup

```{r, message = FALSE}
library(raster) # DEM for map
library(tidyverse) # data processing and plotting functions
library(sf) # tidy spatial data and mapping
library(mgcv) # fit GAMs and GAMMs
library(archdata) # source for the Oxford Pots data
library(tidygraph) # tidy network processing
# devtools::install_github('thomasp85/ggraph') # need dev version of ggraph 
library(ggraph) # network plots
# devtools::install_github('nspope/corMLPE')
library(corMLPE) # network correlation structure for undirected networks
# library(ggmap) # for geocoding, but not used by default (see below)
library(ggspatial) # scale bar for maps
library(maps) # country outlines for maps
```

# Case Studies

## Oxford Pots
First, we'll look at a dataset of Late Romano-British pottery available in the `archdata` package. \emph{\small{David L. Carlson and Georg Roth (2018). archdata: Example Datasets from Archaeological Research. R package version 1.2.}}
```{r}
data("OxfordPots")
```

### Preprocessing
Use `tidyr` and `dplyr` to get the data into a tidy format.
```{r}
pots <- OxfordPots %>%
  as_tibble %>%
  rename(to = Place) %>%
  gather(key, value, OxfordPct:NewForestDst) %>%
  separate(key, c('from', 'measure'), sep = -3) %>%
  spread(measure, value) %>%
  rename(percent = Pct, distance = Dst) %>%
  mutate(from = str_replace(from, 'NewForest', 'New Forest'),
         water = as.factor(WaterTrans == 1)) %>%
  select(to, from, percent, distance, water) %>%
  filter(!is.na(percent))
```

```{r, echo = FALSE}
pots
```

These data don't come with any location information, so we have to do some geocoding using the `ggmap` package. This requires you to get and register your own Google Maps API key, so by default this code chunk is not run and we instead use precomputed values. If you do want to run this chunk, follow the instructions in `?ggmap::register_google` and replace the line `<your-api-key-here>` with your API key.

```{r, eval = FALSE}
ggmap::register_google(key = "<your-api-key-here>")

pots_sites <- pots %>%
  select(from, to) %>% 
  gather %>%
  pull(value) %>%
  unique %>% 
  tibble(site = .) %>%
  # adjust some names 
  mutate(search_term = str_replace_all(site, c('Clausentum' = 'Bitterne', 
                                               'Mildenhall' = 'Mildenhall, Wiltshire')),
         search_term = paste0(search_term, ', UK')) %>%
  mutate_geocode(search_term) %>%
  select(-search_term) %>%
  st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>%
  bind_cols(., as_tibble(st_coordinates(.))) %>%
  rename(x = X, y = Y)

saveRDS(pots_sites, 'oxfordpots_locations.RDS')
```

Join the Oxford Pots data to the geocoded locations.
```{r}
pots_sites <- readRDS('data/oxfordpots_locations.RDS') %>%
  left_join(dplyr::select(pots, to), by = c('site' = 'to'))
```

Make a network object using ``tidygraph` for plotting, and an edgelist tibble for modeling.
```{r}
pots_net <- tbl_graph(nodes = pots_sites, edges = pots, directed = TRUE) %E>%
  filter(!is.na(percent)) %>%
  mutate(similarity = percent / 100)

pots_dat <- pots_net %E>%
  mutate(x = .N()$x[to],
         y = .N()$y[to]) %>% 
  as_tibble %>%
  mutate(from = as.factor(from),
         to = as.factor(to))

pots_net
```

```{r include = FALSE}
# UK boundary shapefile for plotting.
uk_boundary <- maps::map('world', region = 'UK',
                    fill = TRUE, plot = FALSE) %>%
  st_as_sf

pots_net %E>%
  arrange(percent) %N>%
ggraph(., layout = 'manual', x = as_tibble(.)$x, y = as_tibble(.)$y) +
  geom_sf(data = uk_boundary, fill = NA) +
  geom_edge_link(aes(colour = percent), edge_width = 2, lineend = 'round') +
  geom_sf(data = filter(pots_sites, site %in% c('Oxford', 'New Forest')), size = 4) +
  scale_edge_color_viridis(guide = 'legend', name = 'Pottery \nfrom origin site',
                           labels = paste0(seq(5, 20, 5), '%'),
                           breaks = seq(5, 20, 5)) +
  annotation_scale(location = 'br', style = 'ticks') +
  coord_sf(datum = NA, xlim = c(-6, 2), ylim = c(50, 53.8)) +
  ggtitle('Ceramic distribution in Roman Britain', 
          'Percent late Romano-British pottery produced in Oxford and New Forest \n \n \n') +
  theme_void()
```

```{r}
knitr::knit_exit()
```

```{r}
ggplot(pots, aes(distance, percent, color = from)) +
  geom_point(aes(shape = as.factor(WaterTrans)))
```



```{r}
pots_sites <- readRDS('oxfordpots_locations.RDS') %>%
  left_join(select(pots, to, WalledArea) %>% group_by(to) %>% 
              summarise(area = mean(WalledArea)), by = c('site' = 'to'))
```

```{r, eval = FALSE}
st_distance(pots_sites) %>%
  `colnames<-`(pots_sites$site) %>%
  as_tibble %>%
  mutate(to = pots_sites$site) %>%
  gather(from, distance, `New Forest`:Wroxeter) %>%
  filter(to > from) %>%
  right_join(dat, by = c('from' = 'origin', 'to' = 'destination')) %>%
  mutate(distance.x = as.numeric(distance.x) / 1000) %>%
  ggplot(aes(distance.y, distance.x, color = from)) +
  geom_point() +
  geom_label(aes(label = to), alpha = .4)
```

```{r}
pots_net <- tbl_graph(nodes = pots_sites, edges = pots) %E>%
  filter(!is.na(percent)) %>%
  mutate(similarity = percent / 100)
```

```{r}
uk_boundary <- maps::map('world', region = 'UK',
                    fill = TRUE, plot = FALSE) %>%
  st_as_sf
```

```{r}
pots_net %E>%
  arrange(percent) %>%
ggraph() +
  geom_sf(data = uk_boundary, fill = NA) +
  geom_edge_link(aes(colour = percent), edge_width = 2, lineend = 'round') +
  geom_sf(data = filter(pots_sites, site %in% c('Oxford', 'New Forest')), size = 4) +
  scale_edge_color_distiller(palette = 'YlGnBu', guide = 'legend', name = 'Percentage of pottery \nfrom origin site') +
  labs(title = 'Ceramic distribution in Roman Britain', subtitle = 'Percentages of late Romano-British pottery produced in Oxford and New Forest') +
  coord_sf(datum = NA, xlim = c(-6, 2), ylim = c(50, 53.7)) +
  theme_void()
```

```{r}
pots_dat <- pots_net %E>%
  mutate(water = as.factor(WaterTrans == 1),
         x = .N()$x[to],
         y = .N()$y[to]) %>% 
  as_tibble %>%
  mutate(from = as.factor(from),
         to = as.factor(to))
```

Example of model selection.
```{r}
pots_models <- c("s(distance) + from",
            "s(distance) + water + from",
            "s(distance, by = water) + water + from") %>%
  paste('similarity ~', .) %>%
  tibble(formula_chr = .) %>%
  mutate(formula = map(formula_chr, as.formula), 
         model = map(formula, ~gam(.x, method = 'ML', family = betar(), data = pots_dat)),
         aic = map_dbl(model, AIC)) %>%
  arrange(aic)

pots_models
```

Refit the best performing model with REML and investigate.
```{r}
m1 <- gam(similarity ~ s(distance, by = water) + water + from,
          method = 'REML',
          select = TRUE,
          family = betar(),
          data = pots_dat)

summary(m1)
plot(m1, trans = plogis)
```
```{r echo = FALSE}
tmp1 <- expand.grid(distance = 0:140, water = c(TRUE, FALSE), from = 2)
predict(m1, tmp1, type = 'response', se.fit = TRUE) %>%
  bind_cols(tmp1) %>%
  mutate(fit = fit * 100,
         lower = fit - 2 * se.fit * 100,
         upper = fit + 2 * se.fit * 100) %>%
  ggplot(aes(distance, fit, group = water)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = water), alpha = .2) +
  geom_line(aes(y = fit, color = water), size = 1.2) +
  theme_minimal() +
  labs(title = 'Estimated distance decay functions', subtitle = 'With and without water transport',
       x = 'Distance (km)', y = 'Percentage of pottery from origin site')
```

```{r}
pots_net %E>%
  mutate(res = residuals(m1, type = 'response') * 100) %>%
  arrange(abs(res)) %>%
ggraph() +
  geom_sf(data = uk_boundary, fill = NA) +
  geom_edge_link(aes(colour = res), edge_width = 2, lineend = 'round') +
  geom_sf(data = filter(pots_sites, site %in% c('Oxford', 'New Forest')), size = 4) +
  scale_edge_color_distiller(palette = 'RdBu', limits = c(-8, 8), labels = paste0(seq(-8, 8, by = 4), '%'), guide = 'legend', name = '') +
  labs(title = 'Residuals from a distance-only interaction model', subtitle = 'Red ties are stronger then expected by distance, blue ties weaker') +
  coord_sf(datum = NA, xlim = c(-6, 2), ylim = c(50, 53.8)) +
  theme_void()
```

## Data Import and Preprocessing

```{r}
chumash <- read.csv('chumash_marriage.csv', row.names = 1) %>%
  as.matrix %>% 
  replace(. == 0, 999) %>% # replace 0 values with 999 temporarily
  as_tbl_graph(directed = FALSE) %E>%
  filter(!edge_is_loop()) %>%
  rename(marriages = weight) %>%
  mutate(marriages = if_else(marriages == 999, 0, marriages))%N>%
  mutate(centrality = centrality_degree(weights = marriages))
```

```{r}
chumash_locs <- read_csv('chumash_villages_locations.csv') %>%
  st_as_sf(coords = c('X', 'Y'), crs = 4326) %>%
  bind_cols(., as_tibble(st_coordinates(.))) %>%
  rename(x = X, y = Y) %>%
  mutate(size = ordered(size, levels = c('hamlet', 'small', 'medium', 'large', 'very large', 'largest')),
         population = case_when(size == 'hamlet' ~ 15,
          size == 'small' ~ 30,
          size == 'medium' ~ 60,
          size == 'large' ~ 120,
          size == 'very large' ~ 200,
          size == 'largest' ~ 300
          )) 

chumash_net <- left_join(chumash, chumash_locs, by = c('name' = 'village')) %E>%
  mutate(distance = st_distance(.N()$geometry[from], .N()$geometry[to], by_element = TRUE),
         distance = as.numeric(distance) / 1000) %>%
  mutate(eco= case_when(.N()$coastal[from] == .N()$coastal[to] & .N()$coastal[from] == TRUE ~ 'Coastal',
           .N()$coastal[from] == .N()$coastal[to] & .N()$coastal[from] == FALSE ~ 'Inland',
           .N()$coastal[from] != .N()$coastal[to] ~ 'Cross Boundary'),
         eco = as.factor(eco),
         eco_same = as.factor(.N()$coastal[from] == .N()$coastal[to]))

ca_boundary <- maps::map('county', region = 'California',
                    fill = TRUE, plot = FALSE) %>%
  st_as_sf

```

```{r}
chumash_net %E>%
  arrange(marriages) %>%
  filter(marriages > 0) %>%
ggraph() +
  geom_edge_link(aes(colour = marriages, alpha= marriages, width = marriages), lineend = 'round') +
  geom_node_point(aes(size = population, color = coastal)) +
  scale_edge_alpha(range = c(.5, 1)) +
  scale_edge_color_distiller(palette = 'YlGnBu') + 
  theme_void() +
  scale_size_area() +
  coord_quickmap() +
  geom_sf(data = ca_boundary, fill = NA) +
  coord_sf(datum = NA, xlim = c(-120.60246, -119.48277), ylim = c(34.38628, 34.74438 ))

```
468 marriages
```{r}

test <- seq(1.10, 1.30, .01) %>%
  map(~gamm(marriages ~ s(distance, by = eco) + eco,
      method = 'REML',
      select = TRUE,
      family =  Tweedie(p = .x, link = 'log'), #quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges'))))

test2 <- seq(1.10, 1.30, .01) %>%
  map(~gamm(marriages ~ s(distance, by = eco_same) + eco_same,
      method = 'REML',
      select = TRUE,
      family =  Tweedie(p = .x, link = 'log'), #quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges'))))
  
map(test, ~(.$lme) %>% AIC)
map(test2, ~(.$lme) %>% AIC) %>% unlist %>% plot

m1 <- gamm(marriages ~ s(distance, by = eco) + eco,
      method = 'REML',
      select = TRUE,
      family =  Tweedie(p = 1.2, link = 'log'), #quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges')))
m2 <- gamm(marriages ~ s(distance, by = eco) + eco,
      method = 'REML',
      select = TRUE,
      family =  quasipoisson(link = 'log'),
      correlation = corMLPE(form = ~from + to),
      data =   as_tibble(activate(chumash_net, 'edges')))
AIC(m2)

plot(m1$gam, trans = exp)
gratia::draw(m1$gam)
summary(m1$gam)
summary(m1$lme)
gam.check(m1$gam)

chumash_net %E>%
  as_tibble %>%
  ggplot(aes(distance, marriages)) +
  geom_point() +
  geom_smooth()
```

```{r echo = FALSE}
tmp1 <- expand.grid(distance = 0:60, eco = c('Coastal', 'Cross Boundary', 'Inland')) #= c('TRUE', 'FALSE'))
predict(m1$gam, tmp1, type = 'response', se.fit = TRUE) %>%
  bind_cols(tmp1) %>%
  mutate(fit = fit,
         lower = fit - 2 * se.fit,
         upper = fit + 2 * se.fit) %>%
  ggplot(aes(distance, fit, group = eco)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = eco), alpha = .2) +
  geom_line(aes(y = fit, color = eco), size = 1.2) +
  theme_classic() +
  labs(title = 'Estimated distance decay functions', subtitle = 'Within and across ecological boundaries',
       x = 'Distance (km)', y = 'Number of marriages')
```

```{r}
tmp2 <- match(predict(m1$gam), predict(m1$lme))

predict(m1$gam) == predict(m1$lme)

chumash_net %E>%
  mutate(pred = predict(m1$lme),
         pred2 = predict(m1$gam, type = 'response'),
         res = marriages - exp(pred)) %>%
  #mutate(res = residuals(m1$gam, type = 'response')) %>%
  arrange(abs(res)) %>%
  filter(marriages != 0) %>%
ggraph() +
  geom_edge_link(aes(colour = res, alpha= res, width = res), lineend = 'round') +
  geom_node_point(aes(size = population, color = coastal)) +
  scale_edge_alpha(range = c(.5, 1)) +
  scale_edge_color_distiller(palette = 'RdBu', limits = c(-10, 10)) + 
  theme_void() +
  scale_size_area() +
  coord_quickmap() +
  geom_sf(data = ca_boundary, fill = NA) +
  coord_sf(datum = NA, xlim = c(-120.60246, -119.48277), ylim = c(34.38628, 34.74438 ))
```

We can look at the spline functions fit by the GAM. These suggest that overland travel is subject to logarithmic distance decay, but water transport seems to have no such constraints. This is a case where we've used penalization as form of model selection. The data showed no signal of a distance effect for water transport, so the penalized spline was reduced to a flat line.

```{r}
m1 <- gamm(qlogis(similarity) ~ s(distance, bs = 'cr') + as.factor(time), 
           data = dat, 
           correlation = corMLPE(form = ~from + to|as.factor(time)), 
           method = 'ML')
m2 <- gamm(qlogis(similarity) ~ s(distance, bs = 'cr', by = as.factor(time)) + as.factor(time), 
           data = dat, 
           correlation = corMLPE(form = ~from + to|as.factor(time)), 
           method = 'ML')
m3 <- gamm(qlogis(similarity) ~ s(distance, bs = 'cr', m = 2) + s(distance, bs = 'cs', by = as.factor(time), m = 1) +
             as.factor(time), 
           data = dat, 
           correlation = corMLPE(form = ~from + to|as.factor(time)), 
           method = 'ML')
AIC(m1$lme, m2$lme, m3$lme)

plot(m1$gam, trans = plogis)
plot(m2$gam, trans = plogis)
plot(m3$gam, trans = plogis)


AIC(m3$lme, m4$lme)
# moderate support for regional effect
summary(m1$gam)
summary(m2$gam)
summary(m3$lme)


ggplot(dat, aes(distance)) + geom_histogram()
ggplot(dat, aes((similarity))) + geom_histogram()
ggplot(dat, aes(qlogis(similarity), fill = time)) + geom_histogram()
sum(min(dat$similarity))

gam.check(m3$gam)
```

```{r}
mod_time <- dat %>%
  group_by(time) %>%
  nest %>%
  mutate(model = map(data, ~gamm(qlogis(similarity) ~ s(distance, bs = 'cr'), 
                                 data = ., correlation = corMLPE(form = ~from + to), method = 'ML')),
         r2 = map_dbl(model, ~summary(.$gam)$r.sq))
mod_time

mod_time$model %>%
  map(~.$gam) %>%
  walk(plot, trans = plogis)
```

# Next Steps
Problems:
symmetrical vs asymmetrical
lots of zeros

Don't like GAMs? Check out:
1. GERGMS
2. AME models
3. BRMS bayesian implementation?
The underlying concepts are similar, they only differ in how you estimate them

```{r, eval = FALSE}
tmp1 <- expand.grid(distance = 0:40, population = 100, eco_same = c('TRUE', 'FALSE'))
predict(m1$gam, tmp1, type = 'response', se.fit = TRUE) %>%
  bind_cols(tmp1) %>%
  mutate(fit = fit,
         lower = fit - 2 * se.fit,
         upper = fit + 2 * se.fit) %>%
  ggplot(aes(distance, fit, group = eco_same)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = eco_same), alpha = .2) +
  geom_line(aes(y = fit, color = eco_same), size = 1.5) +
  scale_color_manual(values = c('#253494', '#41b6c4')) +
  scale_fill_manual(values = c('#253494', '#41b6c4')) +
  theme_classic() +
  labs(x = 'Distance (km)', y = 'Number of marriages')
title = 'Estimated distance decay functions', subtitle = 'Within and across ecological boundaries',
```

```{r, eval = FALSE}
#tmp2 <- match(predict(m1$gam), predict(m1$lme))

#predict(m1$gam) == predict(m1$lme)

chumash_net %E>%
  mutate(pred = predict(m1$lme),
         pred2 = predict(m1$gam, type = 'response'),
         res = marriages - exp(pred)) %>%
  #mutate(res = residuals(m1$gam, type = 'response')) %>%
  arrange(abs(res)) %>%
  filter(marriages != 0) %>%
ggraph() +
  geom_edge_link(aes(colour = res, alpha= res, width = res), lineend = 'round') +
  geom_node_point(aes(size = baptisms, color = coastal)) +
  scale_edge_alpha(range = c(.5, 1)) +
  scale_edge_color_distiller(palette = 'RdBu', limits = c(-8, 8)) + 
  theme_void() +
  scale_size_area() +
  coord_quickmap() +
  geom_sf(data = ca_boundary, fill = NA) +
  coord_sf(datum = NA, xlim = c(-120.60246, -119.48277), ylim = c(34.38628, 34.74438 ))
```

```{r}
pdf(file = 'figures/tensor_3d.pdf')
vis.gam(m2$gam, view = c('distance', 'baptisms'), theta = 45, phi = 15, type = 'response', se = -1, color = 'bw')
dev.off()
```

```{r}
library(raster)
bbox <- st_as_sfc(st_bbox(chumash_villages))

plot(elev)
srtm <- raster::getData('SRTM', lon = st_coordinates(bbox)[1], 
                        lat = st_coordinates(st_centroid(bbox))[2]) %>%
  crop(as(bbox, 'Spatial')) 

raster::getData('GADM', country = 'USA', level = 1) %>%
  crop(as(bbox, 'Spatial')) %>% plot




plot(srtm)
elev <- srtm %>%
  disaggregate(fact = 2, method = 'bilinear') #%>%
  #as.matrix() %>%
  #t

plot(elev)



  
sphere_shade(elev, texture = 'desert') %>%
  plot_3d(heightmap = elev, zscale = 50, water = TRUE, watercolor = 'desert') %>%
  render_snapshot()
  
render_snapshot()
rgl::rgl.clear()

sphere_shade(elev, texture = 'imhof4') %>%
  plot_map

```

```{r}
library(gdistance)
library(smoothr)
library(ggspatial)

bbox <- extent(-120.7, -119.4500, 34.35000, 34.80000)
elev <- raster('~/SRTM_W_250m_TIF/SRTM_W_250m.tif') %>%
  crop(bbox) %>%
  disaggregate(fact = 2, method = 'bilinear')

writeRaster(elev, 'data/santa_barbara_dem.tif', overwrite = TRUE)
elev[is.na(elev)] <- 99999
altDiff <- function(x){abs(x[2] - x[1])}
hd <- transition(elev, altDiff, 16, symm = TRUE)
#Divide the height differences by the horizontal distances between cells, resulting in slopes.
slope_c <- geoCorrection(hd, type = 'c')
# Figure out which cells are adjacent to one another, queen's case.
adj <- adjacent(elev, cells = 1:ncell(elev), directions = 8)
#Use Tobler's hiking function to calculate walking speed from *cognitive* slope.
tobler_adjusted <- function(x){
  4 * exp(-3.5 * 2.15 * x) / 3.6 # 3.6 turns km/h into m/s
}
speed_c <- slope_c
speed_c[adj] <- tobler_adjusted(slope_c[adj])
#Divide by intercell distance again, resulting in the conductance matrix.
conductance_c <- geoCorrection(speed_c, type = 'c')

nn <- chumash_net %>%
  group_by(from) %>% 
  top_n(-4, distance) %>%
  ungroup


pts <- nn %N>% 
  as_tibble() %>%
  dplyr::select(x, y)

path_dat <- nn %E>%
  as_tibble %>%
  group_by(from) %>%
  nest %>%
  mutate(to = map(data, ~c(.$to) %>% unlist)) %>% # turn the "to" column into list of destinations for each origin
  dplyr::select(-data) %>%
  # calculate the shortest paths
  mutate(paths = map2(from, to, ~shortestPath(conductance_c, 
                                    as.matrix(pts)[.x,], as.matrix(pts)[.y,], 
                                    output = 'SpatialLines'), .progress = TRUE))

paths_sf <- path_dat %>% 
  mutate(paths = map(paths, st_as_sf)) %>% 
  pull(paths) %>% 
  do.call(rbind, .) %>%
  bind_cols((path_dat %>% dplyr::select(-paths) %>% unnest), .) %>%
  st_sf %>%
  filter(., st_is_valid(.)) %>%
  smooth(method = 'ksmooth', smoothness = 10)

mar <- chumash_net %E>% as_tibble %>% dplyr::select(from, to, marriages) %>% filter(marriages > 0)
test_spat <- left_join(nn, paths_sf) %>% dplyr::select(from, to, marriages, distance, geometry) %>%
  mutate(marriages = 0)

for(i in 1:nrow(mar)){
  test_spat <- test_spat %>%
  morph(to_shortest_path, mar[[i,1]], mar[[i,2]], mode = 'all', weights = distance) %>%
  mutate(marriages = marriages + mar[[i,3]]) %>%
  unmorph
}

elev2 <-  elev %>% as.data.frame(xy = TRUE, na.rm = TRUE) %>% as_tibble %>% mutate(elev = na_if(SRTM_W_250m, 99999)) %>% filter(!is.na(elev))

test_spat %E>%
  as_tibble %>%
  st_as_sf %>%
  st_set_crs(4326) %>% st_write('data/chumash_paths.shp')

test_spat %E>%
  as_tibble %>%
  st_as_sf %>%
  st_set_crs(4326) %>%
  arrange(marriages) %>%
  ggplot() +
  geom_raster(data = elev2, aes(x, y, fill = elev), show.legend = FALSE) +
  geom_sf(aes(color = marriages), size = 1.2) +
  geom_sf(data = chumash_villages, aes(size = baptisms), show.legend = 'point') +
  scale_size_area(name = 'Baptisms') +
  geom_sf_label(x = -120.2, y = 34.37, label = 'Santa Barbara Channel',  fontface = 'plain') +
  scale_color_viridis_c(name = 'Marriages') +
  annotation_scale(pad_x = unit(4.35, 'cm'), pad_y = unit(.9, 'cm')) +
  scale_fill_viridis_c(option = 'D', alpha = .6) +
  theme_void() +
  coord_sf(datum = NA)
```

