---
#---POSTER SIZE & DEFAULT FONT---#
poster_height: "36in" # height in inches of poster
poster_width: "48in" # width in inches of poster
font_family: "palatino" # choose from typical latex fonts (example: "palatino")
font_size: "30pt" #please see github.com/brentthorne/posterdown for compatible options.

#---TITLE BOX OPTIONS---#
#ESSENTIALS
title: '\fontfamily{phv}\selectfont Generalized additive mixed models for archaeological network data'
author: "Nicolas Gauthier"
affiliation: "School of Human Evolution and Social Change, Arizona State University"
#STYLE & FORMATTING
titlebox_bgcol: "45ADA8"  #Colour of the Title Box background
titlebox_bordercol: "547980" #Colour of the title Box border.
titlebox_shape: "all"
titlebox_borderwidth: "1cm"
title_textcol: "FFFFFF" #colour of title text
author_textcol: "FFFFFF" # Colour of author text
affiliation_textcol: "FFFFFF" # Colour of affiliation text
title_textsize: "Huge"         # Poster title fontsize
author_textsize: "Large"       # Author list font size
affiliation_textsize: "large"  # Affiliation font size
#ADDING LOGOS
logoleft_name: 'Figures/SHESC'
logoleft_width: '5in'
logoleft_xshift: '1in'
logoleft_yshift: '1in'
logoright_name: 'Figures/SHESC'
logoright_width: '5in'
logoright_xshift: '-1in'
logoright_yshift: '1in'

#---POSTER BODY OPTIONS---#
body_bgcol: "ffffff" #colour of the poster main background
body_textsize: "normalsize"    # Size of the main poster body text
body_textcol: "594F4F"#000000" # Colour of main text in the body of poster
column_numbers: 4 # Number of columns that the poster has
column_margins: "0.5in" # Margin spacing for columns
columnline_col: "008080" #colour 
columnline_width: "0pt" #width of line between each column
#SECTION TITLE STYLING
sectitle_textcol: "ffffff" # Colour of the poster section titles
sectitle_bgcol: "547980" # Colour of the section title box
sectitle_bordercol: "547980" # Colour of the border around the section title box.
sectitle_borderwidth: "2mm" # Thicknes of the section title box border
sectitle_boxshape: "uphill" # Changes the shape of the section title box.

#---BIBLIOGRAPHY OPTIONS---#
bibliography: MyLibrary # name of the .bib file used for referencing
bibliography_spacing: 0.8 # sets the multiplier for line spacing of bibliography spacing (between 0 and 1)
bibliography_textsize: "small"  # size of the bibliography text size (handy for one too many references!)

#---OTHER---#
cite_col: "CC0000" #colour of ciation elements
url_col: "008080" #colour of url links
link_col: "008080" #colour of other links within the poster
footnote_textcol: "ffffff" # Colour of footnote text if used
output: posterdown::posterdown_latex
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis',fig.height = 3.5, fig.width = 4.5, out.width = "100%", warning = FALSE, fig.align = 'center', message = FALSE, dev='cairo_pdf')
```


# Introduction

Distance is a fundamental constraint on human social interaction. This basic principle motivates the use of spatial interaction models for estimating flows of people, information, and resources on spatial and social networks. These models have both valid dynamical and statistical interpretations, a key strength well supported by theory and data from geography, economics, ecology, and genetics. To date, archaeologists have primarily relied on the dynamical approach because the idiosyncrasies of archaeological data make the wholesale adoption of statistical approaches from other fields impractical.
\vspace{1cm}
Here, I argue for the use of generalized additive mixed models (GAMMs) for statistical inference on interaction networks in archaeology. GAMMs are a flexible form of regression model well-matched to the complexities of the archaeological record, including non-normal distributions in the form of counts or proportions, non-independent observations with correlated errors, and non-linear functional relationships. Using two case studies -- an ethnographic marriage network and an archaeological assemblage similarity network -- I illustrate how this approach can lead to unbiased parameter estimates and more robust comparisons of competing hypotheses. I conclude by outlining how future empirical efforts can help to refine our thinking about past network dynamics and reveal cross-cultural regularities of human social interaction in the present day.

# Study Site



```{r, out.width="80%", fig.cap='This is a map of Canada, the ggspatial package is great for GIS folks in R! \\label{mymapfig}', cache=TRUE}

```

# Objectives

\lipsum[2]
\lipsum[4]

# Methods



# Results



\vspace{1cm}

```{r}


```

\vspace{2cm}

```{r, out.width='100%', fig.height=4.5, fig.cap='Using ggplot and patchwork to generate a layout of multiple plots in one figure. The iris dataset was used to generate (a) a line graph, (b) a scatterplot, and (c) a boxplot all together!'}


```


```{r, out.width='85%', fig.width=4.5, fig.height=4, fig.cap='Another typical plot using ggplot, this time with a different theme and r code chunk options for fig.width and fig.height.'}

```


```{r, out.width='80%', fig.width=4.5, fig.align='center', echo=TRUE, fig.height=5, fig.cap='Another figure showing how base R plots might look on this poster!'}



```

\lipsum[1-2]

# Next Steps


<!--- Here you can set the size of the citation text as well as remove the "References" section if you choose not to have one for some reason :) -->
\small\printbibliography
